Combinatorial Identities from Lagrange's Interpolation Polynomial

### 2020-11-12 spectral radius of matrix new method/thesis/article111720.tex
1112 correct errors on the meeting 1110

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[subsection]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{remark}[thm]{Remark}
\newtheorem{eg}[thm]{Example}
\newtheorem{conj}[thm]{Conjecture}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{ex}[thm]{Exercise}
\newtheorem{prob}[thm]{Problem}
\newtheorem{exam}[thm]{Example}
\newtheorem{nota}[thm]{Notation}
\newtheorem{rem}[thm]{Remark}
\newtheorem{ques}[thm]{Question}
\newtheorem{pof}[thm]{Proof.}

@@

\begin{defn}
    When $C$ is an $n \times n$ real matrix, the $\textit {spectral radius} $ $\rho(C)$
        of $C$ is defined by 
        $$\rho(C):=\max\{~|\lambda|~\ |~~\lambda\hbox{ is an eigenvalue of $C$}\},$$
    where $|\lambda|$ is the magnitude of complex number $\lambda.$
\end{defn}

\begin{thm}[Perron–Frobenius theorem],\cite{chang}
    If $C$ is nonnegative square matrix, then the spectral radius $\rho(C)$ is an
    eigenvalue of $C$ with a corresponding nonnogative right eigenvector and a
    corresponding nonnegative left eigenvector.
\end{thm}

@@

The following theorem is from [\cite{content}].

@@

\begin{pof}
    Multiplying the nonnegative vector $u$ in theorem~\ref{pre_thm} assumption
    ~\ref{pre_thm_em1}, where $Qu$ is eigenvector of $C'$,  to the right of both terms of
    ~\ref{pre_thm_em1},    
    \begin{equation}\label{pre1}
       PCQu\leq PC'Qu=\lambda'PQu.
    \end{equation}
    Multiplying the nonnegative left eigenvector $v^T$ of $C$ for $\lambda$ in assumption
     ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}), where $v^TP$ is
    left eigenvector of $C$ for $\lambda$, thus we have
    \begin{equation}\label{pre2}
        \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
    \end{equation}
        Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
        $\lambda\leq \lambda'$ and finish the proof of the first part.
        Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
        Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
        $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
        $u_j\not=0.$ Conversely, (\ref{pre0}) implies $$v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j=
         \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so $\lambda=\lambda'$ by (\ref{pre2}).
\end{pof}

\section{Our Method}
We use [n-1] as notation of the set of elements from one to n-1, which is {1,2,$...$,n}.
Throughout fix $k\in [n-1]$.Let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$ appearing in the position $k,n$ of $E_{kn}$. We will apply the previous theorem with $P=I$ and

@@

\begin{defn}[$k$-rooted vector]
~A column vector $v'=(v'_1,v'_2,\ldots,v'_n)^T$ is called {\it $k$-rooted}  if $v'_{j} \geq 0$ for $1 \leq  j \leq n$ and $v'_k\geq v'_n.$
\end{defn}
\bigskip

The following Lemma is immediate from the above definition.
\bigskip

\begin{lem}[vector rooted lemma]\label{lem:rt_vec}
If $u=(u_1, u_2, \ldots, u_n)^T$ and $v'=(v'_1, v'_2, \ldots, v'_n):=Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$, then
\begin{enumerate}[label=(\roman*)]
\item \label{lem:rt_vec:en1}$v'$ is $k$-rooted  if and only if  $u$ is nonnegative;
\item $u_k>0$ if and only if $v'_k>v'_n$.
\end{enumerate}
\qed
\end{lem}

\begin{thm}[Our first result]\label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\roman*)]
\item \label{thm_main:condition_i} $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item \label{thm_main:condition_ii} there exists a $k$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item \label{thm_main:condition_iii}there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item \label{thm_main:condition_iv}$v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}[label=(\alph*)]
    \item \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
    \item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $with $v_i\ne 0 $;
    \item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$ 
\end{enumerate} \qed
\end{thm}
%

@@

\begin{pof}
    The proof is based on Theorem~\ref{pre_thm} with $P = I$ and $Q = I + E_{kn}$ in (\ref{Q_1} ). 
    The assumption \ref{pre_thm_em1} $PCQ\leq PC'Q$ of Theorem~\ref{pre_thm} holds by the condition \ref{thm_main:condition_i} of this theorem. 
    Let $u = Q^{-1}v'$. Then u is nonnegative and $C'Qu = \lambda' Qu$ by the condition \ref{thm_main:condition_ii} and
     Lemma~\ref{lem:rt_vec}\ref{lem:rt_vec:en1}. Hence the assumption \ref{pre_thm_em2} of Theorem~\ref{pre_thm} holds. The assumptions \ref{pre_thm_em3} and \ref{pre_thm_em4}
      of Theorem~\ref{pre_thm} clearly hold by conditions~\ref{thm_main:condition_iii},\ref{thm_main:condition_iv} of this theorem since $P = I$ and
       $v'= Qu$  Hence $\lambda \leq \lambda' $ by the necessary condition of Theorem~\ref{pre_thm}. Moreover
        $\lambda = \lambda'$ if and only if \ref{pre0} holds, and this is equivalent to
         conditions \ref{thm_main:equ_cond_a},\ref{thm_main:equ_cond_b},\ref{thm_main:equ_cond_c} of this theorem. 
\end{pof}

We are interested in the matrices $C'$ that have $k$-rooted eigenvectors.
Motivated by the condition (i) of theorem 2.3, we provide the following two definitions.

\begin{defn}[(k,n)-sum]
    For an $n \times n$ matrix $C'=(c'_{ij})$, the $(k, n)-sum$ vector of $C'$ is the vector of the sum of the $k$-th and  $n$-th columns of $C'$, where $k\leq n-1$.
\end{defn}

Note that the last column of $C'Q$ is the $(k, n)-sum$ vector of $C'$

\begin{defn}[k-rooted matrix]\label{m_rooted}
    A  matrix $C'=(c'_{ij})$ is called $k-rooted$  if its  columns and its $(k, n)-sum$ vector are all $k-rooted$ except the last column of $C'$.
\end{defn}

%

@@

\begin{rem}
        (i) is immediate from Definition~\ref{m_rooted} and the observation that   
        $$Q^{-1}=I-E_{kn}=\begin{pmatrix}
        1 &  & & &  & 0 \\
         & 1 &  &      &  &  \\
         &  & \ddots & &  & -1 \\
         &  &        & &  &  \\
          &  & & & 1 &  \\
        0 &  & & &  & 1 \\
        \end{pmatrix},$$
\end{rem}
\end{comment}

    \begin{pof}
        and $Q^{-1}C'Q$ is
        $$\begin{pmatrix}
        c'_{11}     & c'_{12} & \cdots     & c'_{1\ n-1} & c'_{1k}+c'_{1n} \\
        \vdots \\
        c'_{k-11}     & c'_{k-1 2}           & \cdots     & c'_{k-1 n-1} & c'_{k-1k}+c'_{k-1n} \\
        c'_{k1}-c'_{n1} & c'_{k2}-c'_{n2} &\cdots      &c'_{kn-1}-c'_{nk-1}& c'_{kk}+c'_{kn}-c'_{nk}-x'_{nn}\\
        c'_{k+11}     & c'_{k+12}           & \cdots     & c'_{k+1\ n-1} & c'_{k+1k}+c'_{k+1n} \\
        \vdots              & \vdots & \ddots              & \vdots & \vdots \\
        c'_{n1}             & c'_{n2} & \cdots             & c'_{n\ n-1} & c'_{nk}+c'_{nn} \\
        \end{pmatrix}.$$
    \end{pof}

    \begin{pof}
        (ii)
            By Lemma~\ref{lem:rt_vec} $v'=Qu$ is $k$-rooted.  
            Since $Q^{-1}C'Qu=\rho(C')u$ by the assumption, we have
            $Q^{-1} C' Q u  = Q^{-1} \rho(C') Qu  =\rho(C')u$  \\
            $C'Qu=\rho(C')Qu$.

@@

    \end{pof}

@@

\begin{pof}
    These assumptions are necessary that $PCQ \leq PC'Q$, and C' is $k$-rooted,based on
    (\ref{lma_m_rooted});\\
    calculation:  \\
       \\
    by \ref{l_diag}, For certain d, if $C'+dI$ is $k$-rooted, then it has
     a $k$-rooted eigenvector with its spectral radius $\lambda + d$. C' would share the same
      eigenvector with $C'+dI$ and has eigenvalue $\lambda$. So $C'+dI$ and $C+dI$ meet the
       conditions of \ref{thm_main}, and we can show that $\rho(C' + dI) \geq \rho(C +dI)$ and
        then $\rho(C') \geq \rho(C)$  \qed
\end{pof}

@@

\section{References}
\begin{thebibliography}{20}
\normalsize
\addcontentsline{toc}{chapter}{Bibliography}

@@

\bibitem{content}
Just as before, {\it What a book}, Fuckes, 2020.

>>>>

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{remark}[thm]{Remark}
\newtheorem{eg}[thm]{Example}
\newtheorem{conj}[thm]{Conjecture}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{ex}[thm]{Exercise}
\newtheorem{prob}[thm]{Problem}
\newtheorem{exam}[thm]{Example}
\newtheorem{nota}[thm]{Notation}
\newtheorem{rem}[thm]{Remark}
\newtheorem{ques}[thm]{Question}
% \newtheorem{pof}[thm]{Proof.}

@@

\begin{defn} \cite{prn_fros1}
    When $C$ is an $n \times n$ real matrix, the $\textit {spectral radius} $ $\rho(C)$
        of $C$ is defined by 
        $$\rho(C):=\max\{~|\lambda|~\ |~~\lambda\hbox{ is an eigenvalue of $C$}\},$$
    where $|\lambda|$ is the magnitude of complex number $\lambda.$
\end{defn}

Perron–Frobenius theorem
\begin{thm} \cite{chang} \cite{prn_fros2}
    If $C$ is nonnegative square matrix, then the spectral radius $\rho(C)$ is an
    eigenvalue of $C$ with a corresponding nonnogative right eigenvector and a
    corresponding nonnegative left eigenvector.
\end{thm}

@@

The following theorem is from \cite{chang}.

@@

\begin{proof}
    Multiplying the nonnegative vector $u$ in theorem~\ref{pre_thm} assumption
    ~\ref{pre_thm_em1}, where $Qu$ is eigenvector of $C'$,  to the right of both terms of
    ~\ref{pre_thm_em1},    
    \begin{equation}\label{pre1}
       PCQu\leq PC'Qu=\lambda'PQu.
    \end{equation}
    Multiplying the nonnegative left eigenvector $v^T$ of $C$ for $\lambda$ in assumption
     ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}), where $v^TP$ is
    left eigenvector of $C$ for $\lambda$, thus we have
    \begin{equation}\label{pre2}
        \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
    \end{equation}
        Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
        $\lambda\leq \lambda'$ and finish the proof of the first part.
        Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
        Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
        $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
        $u_j\not=0.$ Conversely, (\ref{pre0}) implies $$v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j=
         \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so $\lambda=\lambda'$ by (\ref{pre2}).
\end{proof}

\section{Our Method}
We use [n-1] as notation of the set of elements from one to n-1, which is {1,2,$...$,n}.
Throughout fix $k\in [n-1]$.Let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$ appearing in the position $k,n$ of $E_{kn}$. We will apply the previous theorem \ref{pre_thm} with $P=I$ and

@@

\begin{defn}%[$k$-rooted vector]
~A column vector $v'=(v'_1,v'_2,\ldots,v'_n)^T$ is called {\it $k$-rooted}  if $v'_{j} \geq 0$ for $1 \leq  j \leq n$ and $v'_k\geq v'_n.$
\end{defn}
\bigskip

The following Lemma is immediate from the above definition.%[vector rooted lemma]
\bigskip

\begin{lem}\label{lem:rt_vec}
If $u=(u_1, u_2, \ldots, u_n)^T$ and $v'=(v'_1, v'_2, \ldots, v'_n):=Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$, then
\begin{enumerate}[label=(\roman*)]
\item \label{lem:rt_vec:en1}$v'$ is $k$-rooted  if and only if  $u$ is nonnegative;
\item $u_k>0$ if and only if $v'_k>v'_n$.
\end{enumerate}
% \qed
\end{lem}

Below is our first result, in which the first condition implies the first n-1 columns of C major to
 the columns of C', and the (k,n)-sum column of C is also major to C'. The second and the third condition
 suggest that C and C' have nonnegative eigenvectors which are k-rooted. And the forth condition is simplier
  but with the same meaning with theorem (\ref{pre_thm})
\begin{thm}\label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\roman*)]
\item \label{thm_main:condition_i} $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item \label{thm_main:condition_ii} there exists a $k$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item \label{thm_main:condition_iii}there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item \label{thm_main:condition_iv}$v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}[label=(\alph*)]
    \item \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
    \item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $with $v_i\ne 0 $;
    \item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$ 
\end{enumerate} %\qed
\end{thm}
%

@@

\begin{proof}
    The proof is based on Theorem~\ref{pre_thm} with $P = I$ and $Q = I + E_{kn}$ in (\ref{Q_1}). 
    The assumption \ref{pre_thm_em1} $PCQ\leq PC'Q$ of Theorem~\ref{pre_thm} holds by the condition \ref{thm_main:condition_i} of this theorem. 
    Let $u = Q^{-1}v'$. Then u is nonnegative and $C'Qu = \lambda' Qu$ by the condition \ref{thm_main:condition_ii} and
     Lemma~\ref{lem:rt_vec}\ref{lem:rt_vec:en1}. Hence the assumption \ref{pre_thm_em2} of Theorem~\ref{pre_thm} holds. The assumptions \ref{pre_thm_em3} and \ref{pre_thm_em4}
      of Theorem~\ref{pre_thm} clearly hold by conditions~\ref{thm_main:condition_iii},\ref{thm_main:condition_iv} of this theorem since $P = I$ and
       $v'= Qu$  Hence $\lambda \leq \lambda' $ by the necessary condition of Theorem~\ref{pre_thm}. Moreover
        $\lambda = \lambda'$ if and only if \ref{pre0} holds, and this is equivalent to
         conditions \ref{thm_main:equ_cond_a},\ref{thm_main:equ_cond_b},\ref{thm_main:equ_cond_c} of this theorem. 
\end{proof}

We are interested in the matrices $C'$ that have $k$-rooted eigenvectors.
Motivated by the condition (i) of theorem 2.3, we provide the following two definitions. 
This is the definition of (k,n)-sum.
\begin{defn}
    For an $n \times n$ matrix $C'=(c'_{ij})$, the $(k, n)$-sum vector of $C'$ is the vector of the sum of the $k$-th and  $n$-th columns of $C'$, where $k\leq n-1$.
\end{defn}

Note that the last column of $C'Q$ is the $(k, n)$-sum vector of $C'$
Below is the definition of k-rooted matrix.
\begin{defn}\label{m_rooted}
    A  matrix $C'=(c'_{ij})$ is called $k-rooted$  if its  columns and its $(k, n)$-sum vector are all $k-rooted$ except the last column of $C'$.
\end{defn}

@@

\begin{rem}
        (i) is immediate from Definition~\ref{m_rooted} and the observation that   
        $$Q^{-1}=I-E_{kn}=\begin{pmatrix}
        1 &  & & &  & 0 \\
         & 1 &  &      &  &  \\
         &  & \ddots & &  & -1 \\
         &  &        & &  &  \\
          &  & & & 1 &  \\
        0 &  & & &  & 1 \\
        \end{pmatrix},$$
\end{rem}
\end{comment}   
\begin{proof}
    (i)
    and $Q^{-1}C'Q$ is
    $$\begin{pmatrix}
    c'_{11}     & c'_{12} & \cdots     & c'_{1\ n-1} & c'_{1k}+c'_{1n} \\
    \vdots \\
    c'_{k-11}     & c'_{k-1 2}           & \cdots     & c'_{k-1 n-1} & c'_{k-1k}+c'_{k-1n} \\
    c'_{k1}-c'_{n1} & c'_{k2}-c'_{n2} &\cdots      &c'_{kn-1}-c'_{nk-1}& c'_{kk}+c'_{kn}-c'_{nk}-x'_{nn}\\
    c'_{k+11}     & c'_{k+12}           & \cdots     & c'_{k+1\ n-1} & c'_{k+1k}+c'_{k+1n} \\
    \vdots              & \vdots & \ddots              & \vdots & \vdots \\
    c'_{n1}             & c'_{n2} & \cdots             & c'_{n\ n-1} & c'_{nk}+c'_{nn} \\
    \end{pmatrix}.$$
    (ii)
        By Lemma~\ref{lem:rt_vec} $v'=Qu$ is $k$-rooted.  
        Since $Q^{-1}C'Qu=\rho(C')u$ by the assumption, we have
        $Q^{-1} C' Q u  = Q^{-1} \rho(C') Qu  =\rho(C')u$  \\
        $C'Qu=\rho(C')Qu$.

@@

\end{proof}

@@

\begin{proof}
    These assumptions are necessary that $PCQ \leq PC'Q$, and C' is $k$-rooted,based on
    (\ref{lma_m_rooted});\\
    calculation:  \\
       \\
    by \ref{l_diag}, For certain d, if $C'+dI$ is $k$-rooted, then it has
     a $k$-rooted eigenvector with its spectral radius $\lambda + d$. C' would share the same
      eigenvector with $C'+dI$ and has eigenvalue $\lambda$. So $C'+dI$ and $C+dI$ meet the
       conditions of \ref{thm_main}, and we can show that $\rho(C' + dI) \geq \rho(C +dI)$ and
        then $\rho(C') \geq \rho(C)$  %\qed
\end{proof}

@@

%\section*{References}
\begin{thebibliography}{20}
\normalsize
\addcontentsline{toc}{chapter}{Bibliography}

@@
